An Analysis of Security Vulnerabilities in Container Images for Scientific Data Analysis

### 2020-09-14 paper.tex
Edited abstract and intro

\author[1]{Bhupinder Kaur}
\author[1]{Mathieu Dugr\'e}
\author[1]{Aiman Hanna}
\author[1]{Tristan Glatard}

\affil[1]{Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada}

@@

\begin{abstract}
Software containers greatly facilitate the deployment of scientific
data analyses on high-performance computing (HPC) infrastructures, due
to their positive impact on the portability, deployment, and
reproducibility of computations. However, container images often
contain outdated or unnecessary software packages, which increases the
number of security vulnerabilities present in the images, and widens
the attack surface for the entire infrastructure. This paper presents a
vulnerability analysis of container images for scientific data
analysis. We analyze container images with four vulnerability scanners,
focusing on the use case of neuroscience data analyses, and quantifying
the effect of images updates and minification on the number of
vulnerabilities. We conclude with recommendations on how to build
container images with a reduced amount of vulnerabilities.

@@

% Containers are popular on HPC, in particular Singularity
Software containers have emerged has an efficient solution to deploy
scientific data analyses on High Performance Computing (HPC) clusters, due
to their portability, ease of use, and limited overhead. The
Singularity~\cite{kurtzer2017singularity} framework is often preferred to
Docker\footnote{\url{http://docker.com}} for HPC deployments, due to its
secure handling of multi-user environments, and its convenient support for
Docker images. Singularity is now available in dozens of HPC
clusters around the globe, and used routinely for Big Data analyses.

% Containers are meant to isolate applications from the host, but security
% remains an issue
Leveraging core features of the Linux kernel such as namespaces, control
groups and chroot, containers isolate processes from the host computer,  
and control the memory, CPU, network and file-system resources assigned to
them. Nevertheless, containers still share the kernel, mounted file systems
and some devices with the host, which raises security
issues~\cite{martin2018docker, sultan2019container, combe2016docker} and
opens the door to privilege escalation, denial of service, information
leak and other types of attacks~\cite{gantikow2016providing}.

% Container images, our study
Container images often include nearly complete operating-system
distributions in addition to the data analysis software and its required
dependencies. Images are rarely updated, as they are often built once for
the entire duration of an analysis, to avoid results perturbations due to
software updates~\cite{gronenschild2012effects,
glatard2015reproducibility}. Images also commonly include more dependencies
than required by a specific analysis, to facilitate reusability across
experiments. As a result, it has been shown that over 30\% of official
images in DockerHub contain high priority security
vulnerabilities~\cite{gummaraju2015over}, that both official and community
images contain more than 180 vulnerabilities on average~\cite{Shu2017}, and
that vulnerabilities are often due to outdated
packages~\cite{zerouali2019relation}.

In this study, we focus on the vulnerability analysis of container images
used in HPC clusters, focussing on the following questions:

\textit{What is the current amount of known vulnerabilities in
container images deployed on HPC clusters?} Vulnerabilities are possible
attack vectors that may seriously compromise the security of HPC clusters
and the integrity of user data. We report vulnerability analyses produced
by four popular image scanning tools, Anchore, Vuls, Clair, and Stools.

\textit{Can the amount of vulnerabilities be reduced by updating the images?}  
For many reasons related to reproducibility and the life-cycle of research
projects, container images deployed on HPC clusters often include outdated
software. We report on the effect of software updates on the vulnerability
analysis of RQ1.

\textit{Can the amount of vulnerabilities be reduced by minifying the images?} 
Container images often include more software packages than required for a
typical analysis. We report on the impact of this additional software on
the presence of vulnerabilities.

We answer these questions by taking the particular example of neuroimaging,
a domain that has entered the Big Data era as many other
ones~\cite{van2014human}. The next section describes the materials and
tools used in the experiments, including container images and image
scanners. We then describe our image update and minification methodology to
reduce vulnerabilities. Results present the detected vulnerabilities in
container images, explain observed discrepancies between scanners, and
report on the efficiency of image update and size reduction. As a
conclusion, we provide a set of image creation guidelines for a more secure
deployment on HPC clusters.

% We focus on the analysis of container images
% present in BIDS apps~\cite{gorgolewski2017bids} and Boutiques~\cite{glatard2018boutiques}, two containerized
% application frameworks commonly used in neuroimaging.

@@

% \begin{table*}
% 	\csvreader[%
% 	 tabular={|c|c|c|c|c|c|},
% 		table head = \hline\textbf{Labels} &\textbf{{Image names}} & \textbf{Anchore} & \textbf{Clair} & \textbf{Vuls} & \textbf{Stools}\\\hline,
% 		late after line= \\,
% 			late after last line=\\\hline %
% 		]{resul.csv}{labels=\labels,names=\names,A=\A,C=\C,V=\V,Stools=\Stools}%
% 			{\labels & \names & \A & \C & \V & \Stools}
% 		   \centering
% 		\vspace*{1mm}
% 		\caption{\label{table1}Number of vulnerabilities detected by the four scanners. Vuls shows \textit{no data} for images \texttt{bids/antscorticalthickness:v2.2.0-1} and \texttt{bids/afni\_proc:latest}
% 		since their OS distributions reached end of life.} 
% 	\end{table*}
	\begin{figure}
        \includegraphics[width=\columnwidth]{Figures/venn.pdf}
	\caption{\label{fig:venn} Discrepancies in the vulnerabilities detected
	by the different scanners. The Jaccard coefficients between the sets of
	detected vulnerabilities are quite low, showing important discrepancies
	between the scanners: Jaccard(Anchor, Clair) = 0.59, Jaccard(Anchore, Vuls) =
	0.59, Jaccard(Vuls, Clair) = 0.82. Two Ubunty 17.04 images weren't included in this comparison as they cannot be scanned by Vuls. \todo{add Centos images}}
\end{figure}

>>>>

\author{Bhupinder Kaur}
\author{Mathieu Dugr\'e}
\author{Aiman Hanna}
\author{Tristan Glatard}

\affil{Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada}

@@

\begin{abstract}
Software containers greatly facilitate the deployment and reproducibility
of scientific data analyses on high-performance computing clusters
(HPC). However, container images often contain outdated or unnecessary
software packages, which increases the number of security vulnerabilities
in the images and widens the attack surface of the infrastructure. This
paper presents a vulnerability analysis of container images for scientific
data analysis. We analyze container images with four vulnerability
scanners, focusing on the use case of neuroscience data analyses, and
quantifying the effect of images updates and minification on the number of
vulnerabilities. We find that container images used for data analysis
contain hundreds of vulnerabilities, that software updates remove about two
thirds of these vulnerabilities, and that removing unused packages is also
effective. We conclude with recommendations on how to build container
images with a reduced amount of vulnerabilities.

@@

Software containers have emerged has an efficient solution to deploy
scientific data analyses on HPC clusters, due
to their portability, ease of use, and limited overhead. On HPC systems, the
Singularity~\cite{kurtzer2017singularity} framework is often preferred to
\href{http://docker.com}{Docker} because of its
secure handling of multi-user environments and convenient support for
Docker images. Singularity is now available in dozens of HPC
clusters around the world and routinely used for Big Data analysis.

Taking advantage of core Linux kernel features such as namespaces, control
groups and chroot, containers isolate processes from the host computer, and
control the memory, CPU, network and file-system resources assigned to
them. However, containers still share the kernel, mounted file systems and
some devices with the host, which raises security
concerns~\cite{martin2018docker, sultan2019container, combe2016docker} and
opens the door to privilege escalation, denial of service, information leak
and other types of attacks~\cite{gantikow2016providing}.

Container images typically include full operating system distributions in
addition to data analysis software and their dependencies. They are rarely
updated because there is concern that software updates will interfere with
the results~\cite{gronenschild2012effects, glatard2015reproducibility}.
Images also typically include more dependencies than required, to make them
easier to reuse between experiments. As a result, over 30\% of official
images in DockerHub have been shown to contain high priority security
vulnerabilities~\cite{gummaraju2015over}, images on average contain over
180 vulnerabilities~\cite{Shu2017}, and vulnerabilities are often caused by
outdated packages~\cite{zerouali2019relation}.

In this study, we focus on the vulnerabilities present in container images
deployed on HPC clusters for scientific data analysis, and we answer the
following questions:

\textit{What is the current amount of vulnerabilities in
container images deployed on HPC clusters?} Vulnerabilities are possible
attack vectors that can seriously compromise the security of HPC clusters
and the integrity of user data. We report vulnerability scans produced
by four popular image scanning tools: Anchore, Vuls, Clair, and Stools.

\textit{Can the amount of vulnerabilities be reduced by updating the images?}  
For many reasons related to reproducibility and lifecycle of research
projects, container images deployed on HPC clusters often include outdated
software. We report on the effect of software updates on the amount of
vulnerabilities found in images.

\textit{Can the amount of vulnerabilities be reduced by minifying images?} 
Container images often include more software packages than necessary for 
typical analysis. We report on the impact of unused software packages on
the presence of vulnerabilities.

We answer these questions through the example of neuroimaging, a field that
has entered the era of Big Data as many others~\cite{van2014human}. The
following section describes the materials and tools used in the
experiments, including container images and image scanners. We then
describe our methodology for updating and minifying images to reduce
vulnerabilities. The results present the vulnerabilities detected in container
images, quantify the effectiveness of updating and minifying images, and
explain the differences observed between scanners. In conclusion, we
provide a set of image creation guidelines for a more secure deployment of
containers on HPC clusters.

@@

% \begin{table*}
% 	\csvreader[%
% 	 tabular={|c|c|c|c|c|c|},
% 		table head = \hline\textbf{Labels} &\textbf{{Image names}} & \textbf{Anchore} & \textbf{Clair} & \textbf{Vuls} & \textbf{Stools}\\\hline,
% 		late after line= \\,
% 			late after last line=\\\hline %
% 		]{resul.csv}{labels=\labels,names=\names,A=\A,C=\C,V=\V,Stools=\Stools}%
% 			{\labels & \names & \A & \C & \V & \Stools}
% 		   \centering
% 		\vspace*{1mm}
% 		\caption{\label{table1}Number of vulnerabilities detected by the four scanners. Vuls shows \textit{no data} for images \texttt{bids/antscorticalthickness:v2.2.0-1} and \texttt{bids/afni\_proc:latest}
% 		since their OS distributions reached end of life.} 
% 	\end{table*}
	\begin{figure}
        \includegraphics[width=\columnwidth]{Figures/venn.pdf}
	\caption{\label{fig:venn} Discrepancies in the vulnerabilities detected
	by the different scanners. The Jaccard coefficients between the sets of
	detected vulnerabilities are quite low, showing important discrepancies
	between the scanners: Jaccard(Anchor, Clair) = 0.59, Jaccard(Anchore, Vuls) =
	0.59, Jaccard(Vuls, Clair) = 0.82. Two Ubuntu 17.04 images weren't included in this comparison as they cannot be scanned by Vuls. \todo{add Centos images}}
\end{figure}
